PhD Research Proposal: Molecular Level Design of Nanoporous Lyotropic Liquid Crystal Membranes for Aqueous Separations \\ \vspace{0.5cm} \large Advisors: Michael Shirts and Richard Noble

### 2019-05-06 Ben_Manuscripts/comps/Outline.tex
notes.
Fixes in response to comments

\author{Benjamin J. Coscia}

@@

  % re-word. This paragraph is verbatim from structure paper
  More highly selective nanoporous membranes would be useful for 
  performing complex aqueous separations with seawater and various
  types of wastewater.
  \begin{itemize}
    \item For example, sodium chloride and boron in seawater 
    \cite{fritzmann_state---art_2007} and organic micropollutants found in
    municipal and industrial wastewaters \cite{schwarzenbach_challenge_2006}
    represent just a few of the diverse contaminants of water sources. 
    \item By efficiently separating contaminants from feed solutions with
    highly selective membranes, it is possible to reduce the number of 
    required membrane passes and post-treatment steps needed for a given 
    filtration process \cite{werber_materials_2016}, thus lowering cost
    and energy requirements. 
    \item Additionally, one can extract valuable resources from the 
    feed streams. For example, flowback water produced during hydraulic
    fracturing of shale formations contains dissolved species such as acetate
    whose extraction has economic value \cite{dischinger_application_2017}.
  \end{itemize}

@@

  A molecular-level understanding of structure and transport in LLC polymer 
  membranes, enabled by molecular dynamics (MD) simulations, can provide 
  guidelines to reduce the large chemical space available to design
  monomers for creation of separation-specific membranes. 
  \begin{itemize}
    \item Using a sufficiently accurate molecular model, we can directly 
    observe transport of solutes within LLC membrane nanopores and infer
    mechanisms. 
    \item Based on this information we can intelligently design new membranes
    by screening new liquid crystal monomer designs.
    \item Our most promising designs can be tested experimentally.
  \end{itemize}

@@

    Useful molecular-level modeling should incorporate a detailed picture 
    of the nanoscopic pore structure, which is crucial to understanding
    the role of monomer structure in solute transport and membrane design.
    We will generate simulated X-ray diffraction (XRD) patterns from MD trajectories
    and compare them to an experimental 2D wide angle X-ray scattering (WAXS)
    spectrum of a Col\textsubscript{h} phase membrane. We will assess the extent
    to which we can apply our understanding to the H\textsubscript{II} phase,
    as well as systems built with alternate monomers.

@@

    We will observe transport of a relatively large set of small polar solutes
    placed within the H\textsubscript{II} membrane nanopores. We will observe
    the time series of each solute's position as well as directly measure the
    physical interactions, such as hydrogen bonding and ion coordination, 
    between solutes and LLC monomers.

@@

    We will combine our general knowledge of the solute transport mechanisms
    with simulation data in order to inform a stochastic model. This model
    should closely reproduce the time series that we observe in our simulations.
    Due to the low computational cost of a stochastic model relative to 
    MD simulations, we will be able to project long timescale transport behavior
    and make well-converged predictions of macroscopic transport properties.

@@

    Over the course of this project, experimental research surrounding
    LLC membranes has shifted nearly all focus towards the Q\textsubscript{I}
    phase due to its more facile synthesis.	Although most of our work has 
    been applied to the H\textsubscript{II} phase, we expect that much of 
    the same analyses can be applied to the Q\textsubscript{I} phase. The 
    biggest challenge will be adapting our techniques to its more complex
    geometry.

@@

  \section{Progress to Date}\label{section:progress}

@@

  \begin{figure}[!htb]
  \centering
  \begin{subfigure}{\textwidth}
  %BJC: probably change this legend to `d = x Angstroms` instead of 
  % disordered/ordered, since I don't have space to talk about ordered / disorderd configurations (requires 
  % nematic order parameter to give a complete justification of the names)
  \includegraphics[width=\textwidth]{regional_density_legend.pdf}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
        \includegraphics[width=1\linewidth]{head_group_density.pdf}
        \caption{Head Groups}
        \label{fig:head_groups_regional_density}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
        \includegraphics[width=1\linewidth]{tails_density.pdf}
        \caption{Tails}
        \label{fig:tails_regional_density}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
        \includegraphics[width=1\linewidth]{sodium_density.pdf}
        \caption{Sodium Ions}
        \label{fig:sodium_regional_density}
  \end{subfigure}
  \caption{In all cases, the component radial distribution functions are similar.
      They exhibit a composition gradient transitioning from the hydrophilic to the hydrophobic
          regions. The biggest differences are at r=0 where noise is higher due to
          decreased sampling. The center of the pore is not hollow, but contains sodium ions and
          head groups, even when the system is solvated. This architecture may impede transport in
          the real system in a chemically-dependent manner.
          The solvated system has a lower density of head groups near the
          pore center which is likely due to the swelling that is necessary in order to fit water
          molecules in the pore region.}~\label{fig:overlaid_densities}
  \end{figure}

@@

  \noindent The solute MSDs are not a monotonic function of solute size.
  \begin{itemize}
    \item In Figure~\ref{fig:all_msds_10wt}, we plotted the average
    MSD of each solute. 
    \item In Figure~\ref{fig:msd_radius_10wt}, we plotted solute
    MSDs against their molecular radius.
    \item We also plotted theoretical curves which illustrate the 
    expected MSD of each solute if they were to travel unhindered.
    \item We ensured that the Stokes-Einstein equation with the
    correction factor of Gierer and Wirtz\cite{gierer_molekulare_1953}
    passed through methanol in order to give an approximate frame of reference. 
    \item Methanol is small and therefore can travel relatively
    unhindered compared to all other solutes.
    \item The correction factor attempts to include the effects of
    microfriction that begin to play a role when solute size becomes
    on the order of solvent size.
    \item The theoretical line can be used as an approximate boundary
    between subdiffusive, Brownian and superdiffusive behavior.
  \end{itemize}

@@

  \noindent \textbf{\large Objective 3:} \textit{\large Create a stochastic model} (\textcolor{blue}{\textbf{In Progress}})

  %BJC: This section will get developed further as I think about it more after the paper is submitted
  \noindent We are in the process of developing the theory required to build a 
  stochastic model.

>>>>

%MRS: overall, really nice start.  You can cut back a little on the detail and emphasize a bit more ``This is what the data means and what the consequnces are.  Need to flesh out objective 3 a more (and objective 4 a bit more). 
\author{Benjamin J. Coscia}

@@

  % re-word. This paragraph is verbatim from structure paper
  More highly selective nanoporous membranes would be useful for 
  performing complex aqueous separations with seawater and various
  types of wastewater.

  %MRS: probably can do a little less introduction of the overall membrane needs, as we want to make sure 
  %MRS: the focus is on the research (thesis, you can include everything).  Can probably hold off that until you get the rest done, then trim as needed for space.
  %
  %MRS: remember the purpose of the introduction to any document is to lay out the rules of engagement; you want 
  %MRS: to raise the questions that the rest of the document will answer.
  % 
  %MRS: try to make sure the application motivation and the research described match up when possible.
  %MRS: if particular applications wouldn't be directly enabled by the research, at least draw to the connection
  %MRS: i.e. talk about the general tools of computation to understand nanostructured membranes, even if the current
  %MRS: membranes under development aren't suitable for the particular separations described.
  %BJC: Something like: The computational techniques developed here are suitable not just for LLC Membranes, but for any nanostructured 
  % membrane material.
  \begin{itemize}
    \item For example, sodium chloride and boron in seawater 
    \cite{fritzmann_state---art_2007} and organic micropollutants found in
    municipal and industrial wastewaters \cite{schwarzenbach_challenge_2006}
    represent just a few of the diverse contaminants of water sources. 
    \item By efficiently separating contaminants from feed solutions with
    highly selective membranes, it is possible to reduce the number of 
    required membrane passes and post-treatment steps needed for a given 
    filtration process \cite{werber_materials_2016}, thus lowering cost
    and energy requirements. 
    \item Additionally, one can extract valuable resources from the 
    feed streams. For example, flowback water produced during hydraulic
    fracturing of shale formations contains dissolved species such as acetate
    whose extraction has economic value \cite{dischinger_application_2017}.
  \end{itemize}

@@

  A molecular-level understanding of structure and transport in LLC polymer 
  membranes, enabled by molecular dynamics (MD) simulations, can provide 
  guidelines to reduce the large chemical space available to design
  monomers for creation of separation-specific membranes. 
  \begin{itemize}
    \item Using a sufficiently accurate molecular model, we can directly 
    observe transport of solutes within LLC membrane nanopores and infer
    mechanisms. 
    \item Based on this information we will have a much greater capability
    intelligently design new membranes by screening new liquid crystal 
    monomer designs.
    \item The principles learned can be implemented and tested experimentally.
  \end{itemize}

@@

    %MRS: probably should talk about this 1) with proper tense (some of these things have been done) and 2) talk about what YOU are doing, not we. I helped design the research, but you carried essentially all of it out! If phrased in future tense, committee will be confused, thinking those are things still to be done.

    A useful molecular-level model should incorporate a detailed picture 
    of the nanoscopic pore structure, which is crucial to understanding
    the role of monomer structure in solute transport and membrane design.
    I have created a maximally consistent structural model by comparing 
    simulated X-ray diffraction (XRD) patterns, generated from MD trajectories,
    to an experimental 2D wide angle X-ray scattering (WAXS) spectrum of a 
    Col\textsubscript{h} phase LLC membrane. I have used this model as an 
    example in order to understand the ease with which we can apply our approach
    to alternate monomers.

%    shown that we achieve a
%    similar pore composition regardless of initial configuration which is 
%    influenced by a number of variables. This implies that our model can
%    
%        We will assess the extent to which
%    we can apply our understanding to the H\textsubscript{II} phase, as well
%    as systems built with alternate monomers.

@@

    I observed the transport of a relatively large set of small polar solutes
    placed within the H\textsubscript{II} phase membrane nanopores. I analyzed
    the time series of each solute's position in addition to directly measuring
    the physical interactions, such as hydrogen bonding and ion coordination, 
    between solutes and LLC monomers.

@@

    I will combine my qualitative knowledge of the solute transport mechanisms
    with simulation data in order to inform a stochastic model. This model
    should closely reproduce the properties of the time series that we observe 
    in our simulations. Due to the low computational cost of a stochastic model
    relative to MD simulations, I will be able to forecast long timescale transport
    behavior and make well-converged predictions of macroscopic transport properties.

@@

    Over the course of this project, experimental research surrounding
    LLC membranes has shifted nearly all focus towards the Q\textsubscript{I}
    phase due to its more facile synthesis.	Although most of my work has 
    been applied to the H\textsubscript{II} phase, much of the same analyses
    can be applied to the Q\textsubscript{I} phase. The biggest challenge will
    be adapting my techniques to its more complex three dimensional geometry.

@@

%MRS: overall, good description of wha has been done.  You can cut back a little if space is an issue.  Give the paper citations with the points. 
  \section{Progress to Date}\label{section:progress}

@@

  \begin{figure}[!htb]
  \centering
  \begin{subfigure}{\textwidth}
  %BJC: probably change this legend to `d = x Angstroms` instead of 
  % disordered/ordered, since I don't have space to talk about ordered / disorderd configurations (requires 
  % nematic order parameter to give a complete justification of the names)
  \includegraphics[width=\textwidth]{regional_density_legend.pdf}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
        \includegraphics[width=1\linewidth]{head_group_density.pdf}
        \caption{Head Groups}
        \label{fig:head_groups_regional_density}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
        \includegraphics[width=1\linewidth]{tails_density.pdf}
        \caption{Tails}
        \label{fig:tails_regional_density}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
        \includegraphics[width=1\linewidth]{sodium_density.pdf}
        \caption{Sodium Ions}
        \label{fig:sodium_regional_density}
  \end{subfigure}
  %MRS: not clear what is meant by ``all cases''
  \caption{In all cases, the component radial distribution functions are similar.
      They exhibit a composition gradient transitioning from the hydrophilic to the hydrophobic
          regions. The biggest differences are at r=0 where noise is higher due to
          decreased sampling. The center of the pore is not hollow, but contains sodium ions and
          head groups, even when the system is solvated. This architecture may impede transport in
          the real system in a chemically-dependent manner.
          The solvated system has a lower density of head groups near the
          pore center which is likely due to the swelling that is necessary in order to fit water
          molecules in the pore region.}~\label{fig:overlaid_densities}
  \end{figure}

@@

  \noindent The solute MSDs are not a monotonic function of solute size.
  \begin{itemize}
    %MRS: Talk about data presentations (figures) in the present tense.  
    \item In Figure~\ref{fig:all_msds_10wt}, we plotted the average
    MSD of each solute. 
    \item In Figure~\ref{fig:msd_radius_10wt}, we plotted solute
    MSDs against their molecular radius.
    \item We also plotted theoretical curves which illustrate the 
    expected MSD of each solute if they were to travel unhindered.
    \item We ensured that the Stokes-Einstein equation with the
    correction factor of Gierer and Wirtz\cite{gierer_molekulare_1953}
    passed through methanol in order to give an approximate frame of reference. 
    \item Methanol is small and therefore can travel relatively
    unhindered compared to all other solutes.
    \item The correction factor attempts to include the effects of
    microfriction that begin to play a role when solute size becomes
    on the order of solvent size.
    \item The theoretical line can be used as an approximate boundary
    between subdiffusive, Brownian and superdiffusive behavior.
  \end{itemize}

@@

  \noindent \textbf{\large Objective 3:} \textit{\large Create a stochastic model} (\textcolor{blue}{\textbf{In Progress}})
  %MRS: yes, needs to be longer.  you can cut back some of the details in objectives 1 and 2. 
  %BJC: This section will get developed further as I think about it more after the paper is submitted
  \noindent We are in the process of developing the theory required to build a 
  stochastic model.
